Constructive Preference Elicitation by Setwise Max-margin Learning

### 2016-01-15 paper/ijcai16.tex
paper: add first figures, fixup text a bit.

\documentclass{article}
\usepackage{ijcai16}
\usepackage{graphicx}
\usepackage{times}
%\usepackage{latexsym}

@@

\paragraph{Notation.} We use boldface letters $\vx$ to indicate vectors, and
calligraphic capital letters $\calX$ for sets. We frequently abbreviate the set
$\{ x^i \}_{i=1}^n$ as $\{ x^i \}$ whenever the range of the index $i$ is clear
from the context, and use $[n]$ as a shorthand for the set $\{1, \ldots, n\}$.
The $\ell_1$ vector norm is denoted as $\|\vx\|_1 := \sum_z |x_z|$.

@@

We further assume that the set of {\em feasible} configurations is expressed as
a conjunction of linear constraints. This is not a severe restriction:
arbitrary propositional constraints can be trivially encoded in this setting,
e.g. the ``Boolean disjunction'' of two binary variables $x_1 \lor x_2$ can be
rewritten as $x_1 + x_2 \ge 1$. The full set of feasible configurations will be
indicated as $\calX_\text{feasible}$.

In adherence to previous work~\cite{guo2010real,viappiani2010optimal}, we model users by means
of linear utility functions: the user's preferences are represented by a weight
vector $\vw\in\bbR^n$, and the utility of a configuration $\vx$ is given by
$\langle \vw, \vx \rangle = \sum_{z=1}^n w_z x_z$. In the following we will
require all weights to be {\em non-negative} and {\em bounded}: the individual
weights $w_z$ must fall in a (constant but otherwise arbitrary) interval
$[w^\bot_z, w^\top_z]$, with $w^\bot_z \ge 0$. Both requirements are quite
natural, and enable the translation of our core optimization problem into a
mixed-integer linear problem (as done in Section~\ref{sec:formulation}).

During learning, the actual weight vector $\vw$ is {\em unknown} to the
learning system, and must be estimated from pairwise choices elicited from the
user. The set of examples consists of pairwise comparisons between a more
preferred configuration $\vy_+$ and a less preferred on $\vy_-$. Each preference
induces a ranking constraint $\langle \vw, \vy_+ - \vy_- \rangle \ge 0$.
Details on the actual query selection strategy will be provided in
Section~\ref{sec:querystrategy}.
\stefano{add indifference}

@@

A PC configurations is defined by seven attributes: computer type (laptop,
desktop, or tower), manufacturer, CPU model, monitor size, RAM amount, storage
size, and price. See Table~\ref{tab:pcdataset} for an overivew of the
attributes. The price attribute is defined as a linear combination of the other
attributes: this is a fair modeling choice, as often the price of a PC is
well approximated by sum of the price of its components plus an offset due to
branding. Compared to the PC dataset in~\cite{guo2010real}, we combined the CPU
type and CPU frequency attributes into a single CPU attribute (e.g. ``PowerPC
G4 @733 MHz``), so to avoid multiplicative terms in the price function.
\stefano{too detailed/handwavy?}

@@

>>>>

\documentclass{article}
\usepackage{ijcai16}
\usepackage{times}
%\usepackage{latexsym}
\usepackage{graphicx}

@@

\paragraph{Notation.} We use boldface letters $\vx$ to indicate vectors, and
calligraphic capital letters $\calX$ for sets. We frequently abbreviate the set
$\{ x^i \}_{i=1}^n$ as $\{ x^i \}$ whenever the range of the index $i$ is clear
from the context, and use $[n]$ as a shorthand for the set $\{1, \ldots, n\}$.
The $\ell_1$ vector norm is denoted by $\|\vx\|_1 := \sum_z |x_z|$.

@@

We further assume that the set of {\em feasible} configurations is expressed as
a conjunction of linear constraints. This is not a severe restriction:
arbitrary propositional constraints can be trivially encoded in this setting,
e.g. the ``Boolean disjunction'' of two binary variables $x_1 \lor x_2$ can be
rewritten as $x_1 + x_2 \ge 1$. The full set of feasible configurations is
denoted by $\calX_\text{feasible}$.

In adherence to previous work~\cite{guo2010real,viappiani2010optimal}, we model users by means
of linear utility functions: the user's preferences are represented by a weight
vector $\vw\in\bbR^n$, and the utility of a configuration $\vx$ is given by
$\langle \vw, \vx \rangle = \sum_{z=1}^n w_z x_z$. In the following we
require all weights to be {\em non-negative} and {\em bounded}: the individual
weights $w_z$ must fall in a (constant but otherwise arbitrary) interval
$[w^\bot_z, w^\top_z]$, with $w^\bot_z \ge 0$. Both requirements are quite
natural, and enable the translation of our core optimization problem into a
mixed-integer linear problem (as done in Section~\ref{sec:formulation}).

During learning, the actual weight vector $\vw$ is {\em unknown} to the
learning system, and must be estimated from pairwise choices elicited from the
user. The set of examples consists of pairwise comparisons between a more
preferred configuration $\vy_+$ and a less preferred one $\vy_-$. Each preference
induces a ranking constraint $\langle \vw, \vy_+ - \vy_- \rangle \ge 0$.
Details on the actual query selection strategy will be provided in
Section~\ref{sec:querystrategy}.
\stefano{add indifference}

@@

A PC configurations is defined by seven attributes: computer type (laptop,
desktop, or tower), manufacturer, CPU model, monitor size, RAM amount, storage
size, and price. See Table~\ref{tab:pcdataset} for an overivew of the
attributes. The price attribute is defined as a linear combination of the other
attributes: this is a fair modeling choice, as often the price of a PC is
well approximated by sum of the price of its components plus an offset due to
branding. Compared to the PC dataset in~\cite{guo2010real}, we combined the CPU
type and CPU frequency attributes into a single CPU attribute (e.g. ``PowerPC
G4 @733 MHz''), so to avoid multiplicative terms in the price function.
\stefano{too detailed/handwavy?}

@@

\onecolumn

\begin{figure}[b]
    \centering
    \begin{tabular}{cccc}
        {\bf synthetic 2} & {\bf synthetic 3} & {\bf synthetic 4} & {\bf synthetic 5}
        \\
        \includegraphics[width=12em]{figure_synthetic_2_uniform} &
        \includegraphics[width=12em]{figure_synthetic_3_uniform} &
        \includegraphics[width=12em]{figure_synthetic_4_uniform} &
        \includegraphics[width=12em]{figure_synthetic_5_uniform}
        \\
        \includegraphics[width=12em]{figure_synthetic_2_uniform_sparse} &
        \includegraphics[width=12em]{figure_synthetic_3_uniform_sparse} &
        \includegraphics[width=12em]{figure_synthetic_4_uniform_sparse} &
        \includegraphics[width=12em]{figure_synthetic_5_uniform_sparse}
        \\
        \includegraphics[width=12em]{figure_synthetic_2_normal} &
        \includegraphics[width=12em]{figure_synthetic_3_normal} &
        \includegraphics[width=12em]{figure_synthetic_4_normal} &
        \includegraphics[width=12em]{figure_synthetic_5_normal}
        \\
        \includegraphics[width=12em]{figure_synthetic_2_normal_sparse} &
        \includegraphics[width=12em]{figure_synthetic_3_normal_sparse} &
        \includegraphics[width=12em]{figure_synthetic_4_normal_sparse} &
        \includegraphics[width=12em]{figure_synthetic_5_normal_sparse}
    \end{tabular}
    \caption{Results on the synthetic dataset. Each row represents a different
    setting: uniform, uniform sparse, normal, normal sparse. Color indicates
    the set size $k$: blue means $k=2$, greed $k=3$, and red $k=4$.}
\end{figure}

\twocolumn
